Rust for Scientific Computing

### 2025-08-22 main.tex
add more notes for introduction

>>>>

General Remarks about languages
\begin{enumerate}
    \item Programming languages allow us to express machine instructions in human-readable code.
    \item Depend on use-case
    \item Translate human-interpretable code into machine instructions
    \item tradeoff between high-level view (abstraction) and performance (low level)
    \item C++ has "zero-cost" abstractions
\end{enumerate}

Importance of programming in science
\begin{enumerate}
    \item New fields of study (computational physics, chemistry, biology, sociology, etc.)
    \item Cite some groundbreaking examples which were possible due to programming in science (maybe
          weather)
    \item
\end{enumerate}

Judging C/C++
\begin{enumerate}
    \item Previous Decades: C/C++ in science
    \item Limitations of design choices of language become evident
    \item fragmented ecosystem; good for competition; bad for developer experience
    \item modern features missing: package manager with package registries, build system
    \item technical details; mutable by default, templates vs generics, backwards compatiblity
          becomes a problem ("there is a much simpler langauge within C++ that tries to get out"),
          macros, no clear ownership rules on a language level, dangling pointers
    \item still many good principles (such as RAII) and design principles/concepts (cite gang of
          four)
\end{enumerate}

The need for modern languages
\begin{enumerate}
    \item Java aimed to replace C due to pointers
    \item Go also aims to replace it, go is garbage-collected; thus cannot be as well-performing as
          C++ or C
    \item Python, Javascript for high-level abstractions ("as glue"); no performance howerver
\end{enumerate}

Elaborate on why Rust solves many problems
\begin{enumerate}
    \item safety guarantees; ensures that code which we write actually performs what we set out to
          do
    \item no dangling pointers
    \item zero-cost abstrations
    \item compiled without garbage-collector
    \item tooling: modern ecosystem; package manager, build system, package registry
    \item strong type system which is generally considered sound
    \item macro system gets type-checked
    \item very good C-interop; utilize existing software
\end{enumerate}

Show that many companies are starting to adopt Rust
\begin{enumerate}
    \item Amazon (AWS)\\
          AWS developed Firecracker, a microVM monitoring service powering AWS Lambda and AWS
          Fargate.
          Firecracker is built primarily in Rust to offer fast cold boot times and secure
          workload isolation.
    \item Rust-supported runtimes like AWS Lambda also allow developers to write serverless
          functions using Rust.
    \item Google\\
          Fuchsia, Google's modern operating system, uses Rust extensively for system components
          such as drivers and services to enhance security and reliability.
    \item Google (Android)\\
          Android integration: Rust has been adopted within the Android Open Source Project (AOSP)
          to improve memory safety in critical system components. DEV Community Litslink
    \item Meta (Facebook)\\
          Mononoke, Facebook’s large-scale source control backend for Mercurial, was written in
          Rust to improve performance and reliability in its monorepo.
    \item Meta (Facebook)\\
          Facebook employs Rust in developing CLI tools, distributed systems, and cron jobs.
          Engineers write millions of lines of Rust internally.
    \item Apple\\
          Apple’s Cloud Traffic team and backend infrastructure have been shifting from C to Rust,
          with new functionality being primarily built in Rust. This change was driven by
          performance, security, and developer efficiency goals. Reddit
    \item Microsoft\\
          Although not part of FAANG, it’s noteworthy—Microsoft is rewriting parts of the Windows
          kernel and system components in Rust (e.g., DWriteCore, GDI Regions), prioritizing memory
          safety.
    \item Azure services-like Azure IoT Edge—also incorporate Rust.
    \item Linux Kernel\\
          Only other supported language except from C
\end{enumerate}
